---
title: "Exploratory Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
---

## Temporal and Spatial Trends in Rodent Complaints

### 1. Temporal Trends

By visualizing the trend of rodent complaints over time, we can see that there was a slight decline in complaints between June 2022 and December 2022. However, complaints increased significantly in June 2023 and continued to rise through December 2023.


<details>
<summary>Click to view the detailed code</summary>

```{r temporal-trends, echo=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)

# Load cleaned data
complaints_clean <- read.csv("cleaned_311_complaints.csv")

# Convert `date` to Date type
complaints_clean <- complaints_clean %>%
  mutate(date = as.Date(date))

# Extract temporal features (Year and Month)
complaints_clean <- complaints_clean %>%
  mutate(
    year = year(date),
    month = month(date, label = TRUE)
  )

# Temporal Trends Analysis
temporal_trend <- complaints_clean %>%
  group_by(year, month) %>%
  summarise(total_complaints = sum(count, na.rm = TRUE)) %>%
  arrange(year, month)
```
</details>


```{r}
# Plot temporal trends
ggplot(temporal_trend, aes(x = interaction(year, month, sep = "-"), y = total_complaints, group = year)) +
  geom_line(color = "blue", linewidth = 1) +
  labs(
    title = "Temporal Trends in Rodent Complaints",
    x = "Year-Month",
    y = "Total Complaints"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

### 2. Spatial Trends

Visualize the distribution of rodent complaints across different regions.

<details>
<summary>Click to view the detailed code</summary>

```{r}
# Spatial Trends Analysis
spatial_trend <- complaints_clean %>%
  group_by(region) %>%
  summarise(total_complaints = sum(count, na.rm = TRUE)) %>%
  arrange(desc(total_complaints))
```
</details>

```{r}
# Plot spatial trends
ggplot(spatial_trend, aes(x = reorder(region, total_complaints), y = total_complaints, fill = region)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(
    title = "Spatial Trends in Rodent Complaints by Region",
    x = "Region",
    y = "Total Complaints"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

### 3. Temporal Trends by Region
Analyze complaint trends over time for the regions with the highest complaints.

<details>
<summary>Click to view the detailed code</summary>
```{r}
# Temporal Trends by Region
temporal_spatial_trend <- complaints_clean %>%
  group_by(region, year, month) %>%
  summarise(total_complaints = sum(count, na.rm = TRUE), .groups = "drop") %>%
  arrange(region, year, month)

# Filter for top regions
top_regions <- spatial_trend %>%
  top_n(5, total_complaints) %>%
  pull(region)

temporal_spatial_filtered <- temporal_spatial_trend %>%
  filter(region %in% top_regions)
```
</details>

```{r}
# Plot temporal trends for top regions
ggplot(temporal_spatial_filtered, aes(x = interaction(year, month, sep = "-"), y = total_complaints, color = region, group = region)) +
  geom_line(linewidth = 1) +
  labs(
    title = "Temporal Trends in Rodent Complaints for Top Regions",
    x = "Year-Month",
    y = "Total Complaints"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```



