---
title: "Regression Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
---

## 

### Defining variables
Dependent Variable (Y): Number of complaints (count from the 311 complaints dataset).
Independent Variables (X):
Exterminator visits (count from the exterminator visits dataset).
Compliance inspections (count from the compliance inspections dataset).

Steps:
Merge the datasets based on <code>date</code> and <code>zone_id</code>.
Run a linear regression model to assess how independent variables influence complaints.
Evaluate the model.

<details>
<summary>Click to view the detailed code</summary>

```{r}
library(dplyr)
library(ggplot2)
library(broom)  # For model summary

# Load cleaned datasets
complaints_data <- read.csv("cleaned_311_complaints.csv")
exterminator_data <- read.csv("cleaned_exterminator_visits.csv")
inspections_data <- read.csv("cleaned_compliance_inspections.csv")

# Aggregate complaints data
complaints_data <- complaints_data %>%
  group_by(date, zone_id, region) %>%
  summarise(complaints = sum(count, na.rm = TRUE), .groups = "drop")

# Aggregate exterminator visits data
exterminator_data <- exterminator_data %>%
  group_by(date, zone_id, region) %>%
  summarise(exterminator_visits = sum(count, na.rm = TRUE), .groups = "drop")

# Aggregate compliance inspections data
inspections_data <- inspections_data %>%
  group_by(date, zone_id, region) %>%
  summarise(compliance_inspections = sum(count, na.rm = TRUE), .groups = "drop")

# Merge datasets by date, zone_id, and region
merged_data <- complaints_data %>%
  inner_join(exterminator_data, by = c("date", "zone_id", "region")) %>%
  inner_join(inspections_data, by = c("date", "zone_id", "region"))



```
</details>

```{r}
# Check the structure of the merged dataset
summary(merged_data)

```
<details>
<summary>Click to view the detailed code</summary>
```{r}
# Run Linear Regression: Complaints ~ Exterminator Visits + Compliance Inspections
regression_model <- lm(complaints ~ exterminator_visits + compliance_inspections, data = merged_data)

# Display Model Summary
model_summary <- summary(regression_model)
print(model_summary)

# Extract Regression Coefficients
coefficients <- broom::tidy(regression_model)
print(coefficients)

# Evaluate Model Fit: Extract R-squared and Adjusted R-squared
r_squared <- model_summary$r.squared
adjusted_r_squared <- model_summary$adj.r.squared

# Print R-squared and Adjusted R-squared
cat("R-squared: ", r_squared, "\n")
cat("Adjusted R-squared: ", adjusted_r_squared, "\n")

```
</details>

```{r}
# Visualize Regression Relationships

# Scatter plot with regression line for exterminator visits
ggplot(merged_data, aes(x = exterminator_visits, y = complaints)) +
  geom_point(alpha = 0.6, color = "blue") +
  geom_smooth(method = "lm", color = "red", se = TRUE) +
  labs(
    title = "Relationship Between Exterminator Visits and Complaints",
    x = "Exterminator Visits",
    y = "Number of Complaints"
  ) +
  theme_minimal()

# Scatter plot with regression line for compliance inspections
ggplot(merged_data, aes(x = compliance_inspections, y = complaints)) +
  geom_point(alpha = 0.6, color = "green") +
  geom_smooth(method = "lm", color = "purple", se = TRUE) +
  labs(
    title = "Relationship Between Compliance Inspections and Complaints",
    x = "Compliance Inspections",
    y = "Number of Complaints"
  ) +
  theme_minimal()
```

```{r}
# Fit the regression model
regression_model <- lm(complaints ~ exterminator_visits + compliance_inspections, data = merged_data)

# Display the full model summary
model_summary <- summary(regression_model)

# Extract R-squared and Adjusted R-squared
r_squared <- model_summary$r.squared
adjusted_r_squared <- model_summary$adj.r.squared

# Print the values
cat("R-squared: ", r_squared, "\n")
cat("Adjusted R-squared: ", adjusted_r_squared, "\n")
```

