---
title: "Regression Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
---

### Defining variables
##### Dependent Variable (Y): 
* Number of complaints (from the 311 complaints dataset).

##### Independent Variables (X):
* Exterminator visits (from the exterminator visits dataset).
* Compliance inspections (from the compliance inspections dataset).

Steps:

* Merge the datasets based on <code>date</code> and <code>zone_id</code>.
* Run a linear regression model to assess how independent variables influence complaints.
* Evaluate the model.

<details>
<summary>Click to view the detailed code</summary>

```{r}
library(dplyr)
library(ggplot2)
library(broom)

complaints_data <- read.csv("cleaned_311_complaints.csv")
exterminator_data <- read.csv("cleaned_exterminator_visits.csv")
inspections_data <- read.csv("cleaned_compliance_inspections.csv")

complaints_data <- complaints_data %>%
  group_by(date, zone_id, region) %>%
  summarise(complaints = sum(count, na.rm = TRUE), .groups = "drop")

exterminator_data <- exterminator_data %>%
  group_by(date, zone_id, region) %>%
  summarise(exterminator_visits = sum(count, na.rm = TRUE), .groups = "drop")

inspections_data <- inspections_data %>%
  group_by(date, zone_id, region) %>%
  summarise(compliance_inspections = sum(count, na.rm = TRUE), .groups = "drop")

merged_data <- complaints_data %>%
  inner_join(exterminator_data, by = c("date", "zone_id", "region")) %>%
  inner_join(inspections_data, by = c("date", "zone_id", "region"))

```
</details>

```{r}
# Check the structure of the merged dataset
summary(merged_data)

```
<details>
<summary>Click to view the detailed code</summary>
```{r}
# Run Linear Regression: Complaints ~ Exterminator Visits + Compliance Inspections
regression_model <- lm(complaints ~ exterminator_visits + compliance_inspections, data = merged_data)

# Display Model Summary
model_summary <- summary(regression_model)
print(model_summary)

```
</details>

##### The fitted model:
Complaints=377.12+0.26×Exterminator Visits−0.08×Compliance Inspections
<p><br>

### Coefficient Analysis (from <code>broom::tidy()</code>)
  For every <strong>additional exterminator</strong> visit, complaints are expected to increase by approximately 0.258, when inspections remain constant.
  The p-value is 0.0263, meaning that it is significant at the 5% level of significance.
  This positive relationship suggests that exterminator visits are associated with more complaints, indicating a reactive response, where exterminator visits are higher in areas with more complaints.

  For every additional compliance inspection, complaints are expected to decrease by approximately 0.084, when exterminator visits remain constant.
  The p-value is 0.299, indicating it is not statistically significant at the 5% level of significance.
  The negative estimate suggests a potential reduction in complaints with more inspections, but the lack of statistical significance means this effect cannot be reliably inferred from the data.

<details>
<summary>Click to view the detailed code</summary>

```{r}
# Extract Regression Coefficients
coefficients <- broom::tidy(regression_model)
print(coefficients)
```
</details>
<p><br>

### Model Fit (from <code>summary()</code>)
<p>* R_squared: 0.4046<br>
Approximately 40.46% of the variability in complaints is explained by the model. This is a moderate fit but leaves 59.54% of the variability unexplained, suggesting other factors may contribute to complaints.

<p>* Adjusted R-squared: 0.3130<br>
After accounting for the number of predictors, the model explains 31.3% of the variability in complaints. The relatively lower value of adjusted R-squared indicates that some variance may be attributed to noise or irrelevant predictors.
<p><br>

### Visualize Regression Relationships

##### Exterminator Visits vs. Complaints:
* The scatter plot shows a positive trend with a significant regression line (p < 0.05).
* The shaded confidence interval highlights variability in the predictions, with more uncertainty at higher numbers of visits.

##### Compliance Inspections vs. Complaints:
* The plot shows a slight negative trend, but the regression line is not statistically significant (p > 0.05).
* The confidence interval is relatively wide, reflecting the weak relationship.

<details>
<summary>Click to view the detailed code</summary>
```{r}
# Load necessary libraries
library(plotly)

# Create an interactive scatter plot with regression line for exterminator visits
scatter_exterminator <- plot_ly(
  data = merged_data,
  x = ~exterminator_visits,
  y = ~complaints,
  type = "scatter",
  mode = "markers",
  marker = list(color = "blue", opacity = 0.6)
) %>%
  add_trace(
    x = ~exterminator_visits,
    y = fitted(lm(complaints ~ exterminator_visits, data = merged_data)),
    type = "scatter",
    mode = "lines",
    line = list(color = "red"),
    name = "Regression Line"
  ) %>%
  layout(
    title = "Relationship Between Exterminator Visits and Complaints",
    xaxis = list(title = "Exterminator Visits"),
    yaxis = list(title = "Number of Complaints")
  )

# Display the plot
```
</details>

```{r}
scatter_exterminator

```


<details>
<summary>Click to view the detailed code</summary>

```{r}
scatter_compliance <- plot_ly(
  data = merged_data,
  x = ~compliance_inspections,
  y = ~complaints,
  type = "scatter",
  mode = "markers",
  marker = list(color = "green", opacity = 0.6)
) %>%
  add_trace(
    x = ~compliance_inspections,
    y = fitted(lm(complaints ~ compliance_inspections, data = merged_data)),
    type = "scatter",
    mode = "lines",
    line = list(color = "purple"),
    name = "Regression Line"
  ) %>%
  layout(
    title = "Relationship Between Compliance Inspections and Complaints",
    xaxis = list(title = "Compliance Inspections"),
    yaxis = list(title = "Number of Complaints")
  )
```
</details>

```{r}
scatter_compliance
```

<p><br>


### Adding Aditional Time Predictor

##### Exterminator Visits:
* Estimate: 0.25798, p-value = 0.0329 </p>
* Statistically significant at the 5% level.
* A positive coefficient indicates that for every additional exterminator visit, complaints are expected to increase by approximately 0.258, when all other variables remain constant.

##### Compliance Inspections:
* Estimate: -0.08301, p-value = 0.3259
* Not statistically significant.
* A negative coefficient indicates a potential decrease in complaints with more inspections, but the evidence is not strong enough to draw reliable conclusions.

##### Month:
* Estimate: -5.66314, p-value = 0.9068
* Not statistically significant.
* This suggests that complaints do not show a strong linear trend across months. 

<details>
<summary>Click to view the detailed code</summary>
```{r}
# Extract month as a numeric value and categorize into seasons
merged_data <- merged_data %>%
  mutate(
    month = format(as.Date(date), "%m"),  
    season = case_when(                  
      month %in% c("12", "01", "02") ~ "Winter",
      month %in% c("03", "04", "05") ~ "Spring",
      month %in% c("06", "07", "08") ~ "Summer",
      month %in% c("09", "10", "11") ~ "Fall"
    ),
    month = as.numeric(month)            # Convert month to numeric for regression
  )

# Convert season to a factor for regression analysis
merged_data <- merged_data %>%
  mutate(season = factor(season, levels = c("Winter", "Spring", "Summer", "Fall")))

# Run the updated regression model: Complaints ~ Exterminator Visits + Compliance Inspections + Time Predictors
regression_model_time <- lm(complaints ~ exterminator_visits + compliance_inspections + month + season, data = merged_data)

model_summary_time <- summary(regression_model_time)
print(model_summary_time)

# Extract and print R-squared and Adjusted R-squared
r_squared_time <- model_summary_time$r.squared
adjusted_r_squared_time <- model_summary_time$adj.r.squared
cat("R-squared: ", r_squared_time, "\n")
cat("Adjusted R-squared: ", adjusted_r_squared_time, "\n")

# Extract regression coefficients
coefficients_time <- broom::tidy(regression_model_time)
print(coefficients_time)

```
</details>
<p><br>

The boxplot below shows that complaints are generally higher during Summer than Winter, with wider variability in the Summer. Food availability could be a potential reason for accounting this pattern. Although seasonal effect is not significant in this model, further investigation into weather and food availability for rodents may reveal actionable insights.

<details>
<summary>Click to view the detailed code</summary>
```{r}
library(plotly)

plotly_boxplot <- plot_ly(
  data = merged_data,
  x = ~season,
  y = ~complaints,
  type = "box",
  color = ~season,
  colors = c("Winter" = "blue", "Summer" = "orange")
) %>%
  layout(
    title = "Seasonal Effect on Complaints",
    xaxis = list(title = "Season"),
    yaxis = list(title = "Number of Complaints"),
    showlegend = FALSE
  )

```

</details>

```{r}
plotly_boxplot
```
<p><br>

### Adding Interaction Term <code>exterminator_visits * season</code>

The interaction term is not significant, indicating that the relationship between exterminator visits and complaints does not vary meaningfully between Winter and Summer.

<details>
<summary>Click to view the detailed code</summary>
```{r}
# Update the regression model to include the interaction term
regression_model_interaction <- lm(
  complaints ~ exterminator_visits * season + compliance_inspections + month,
  data = merged_data
)

# Display the updated model summary
model_summary_interaction <- summary(regression_model_interaction)
print(model_summary_interaction)

# Extract and print R-squared and Adjusted R-squared
r_squared_interaction <- model_summary_interaction$r.squared
adjusted_r_squared_interaction <- model_summary_interaction$adj.r.squared
cat("R-squared: ", r_squared_interaction, "\n")
cat("Adjusted R-squared: ", adjusted_r_squared_interaction, "\n")

# Extract regression coefficients
coefficients_interaction <- broom::tidy(regression_model_interaction)
print(coefficients_interaction)

# Create a scatter plot with regression lines for each season
interaction_plot <- plot_ly(
  data = merged_data,
  x = ~exterminator_visits,
  y = ~complaints,
  color = ~season,
  type = "scatter",
  mode = "markers",
  marker = list(opacity = 0.6)
) %>%
  # Add regression lines for each season
  add_trace(
    data = merged_data %>% filter(season == "Winter"),
    x = ~exterminator_visits,
    y = fitted(lm(complaints ~ exterminator_visits, data = merged_data %>% filter(season == "Winter"))),
    type = "scatter",
    mode = "lines",
    line = list(color = "blue", dash = "solid"),
    name = "Winter Regression"
  ) %>%
  add_trace(
    data = merged_data %>% filter(season == "Summer"),
    x = ~exterminator_visits,
    y = fitted(lm(complaints ~ exterminator_visits, data = merged_data %>% filter(season == "Summer"))),
    type = "scatter",
    mode = "lines",
    line = list(color = "orange", dash = "solid"),
    name = "Summer Regression"
  ) %>%
  layout(
    title = "Interaction Between Exterminator Visits and Season",
    xaxis = list(title = "Exterminator Visits"),
    yaxis = list(title = "Number of Complaints"),
    legend = list(title = list(text = "Season"))
  )


```
</details>
``

```{r}
interaction_plot

```
<p><br>

### Quantile Regression

By analyzing quantile regression for this project, we could have a deeper understanding of how exterminator visits and compliance inspections impact rodent complaints across neighborhoods with varying levels of infestations. Unlike traditional regression, which focuses on average effects, it allows us to analyze how these predictors influence areas with low, moderate, and high complaint volumes. The results will reveal whether exterminator visits have a disproportionately higher effect in neighborhoods with severe rodent issues (upper quantiles) compared to areas with minimal complaints (lower quantiles). This insight is essential for NYC’s efforts to combat rodent infestations, as it helps the Department of Sanitation and related agencies tailor their strategies to specific community needs.

<details>
<summary>Click to view the detailed code</summary>
```{r}
library(quantreg)

quantile_model_25 <- rq(complaints ~ exterminator_visits + compliance_inspections, tau = 0.25, data = merged_data)
quantile_model_50 <- rq(complaints ~ exterminator_visits + compliance_inspections, tau = 0.50, data = merged_data) 
quantile_model_75 <- rq(complaints ~ exterminator_visits + compliance_inspections, tau = 0.75, data = merged_data)

cat("Quantile Regression for Tau = 0.25\n")
print(summary(quantile_model_25))

cat("\nQuantile Regression for Tau = 0.50 (Median)\n")
print(summary(quantile_model_50))

cat("\nQuantile Regression for Tau = 0.75\n")
print(summary(quantile_model_75))

# Visualize quantile regression results
taus <- c(0.25, 0.5, 0.75)  # Quantiles to analyze
quantile_results <- rq(complaints ~ exterminator_visits + compliance_inspections, tau = taus, data = merged_data)

# Coefficients across quantiles
quantile_coefficients <- coef(quantile_results)
print(quantile_coefficients)

```
</details>
<p><br>

#### Effect of Exterminator Visits:
The relationship strengthens as complaints increase (from Tau = 0.25 to Tau = 0.75).
At high quantiles (e.g., Tau = 0.75), exterminator visits are significantly associated with complaints. This suggests exterminator visits are primarily deployed reactively in high-complaint areas, rather than proactively preventing complaints.

#### Effect of Compliance Inspections:
The negative relationship between compliance inspections and complaints becomes stronger at higher quantiles, potentially reflecting the efficacy of inspections in high-complaint areas. However, the effects are not statistically significant.

#### Heterogeneity Across Quantiles:
The results show substantial differences in the effects of predictors across quantiles, emphasizing the importance of quantile regression in capturing these variations.

```{r}
par(mar = c(3.9, 3.6, 4, 2) + 0.5)

plot(
  summary(quantile_results),
  parm = "exterminator_visits",
  main = "Effect of Exterminator Visits Across Quantiles"
)

title(
  xlab = "Quantiles of Complaints",
  ylab = "Coefficient for Exterminator Visits"
)


```
<p><br>

### Conclusion
  The quantile regression analysis reveals important insights into how exterminator visits influence rodent complaints across different levels of complaints. The coefficients for exterminator visits increase as we move from lower quantiles (0.25) to higher quantiles (0.75), suggesting that exterminator visits have a stronger impact in areas with a higher volume of complaints. At the 0.25 quantile, the effect is minimal (0.037), while at the 0.75 quantile, the coefficient rises to 0.335. This indicates that neighborhoods with more severe rodent problems (upper quantiles of complaints) are more sensitive to the number of exterminator visits.

  The confidence intervals, as shown in the shaded region of the plot, suggest that while the effect is statistically uncertain at lower quantiles, it becomes more pronounced and meaningful at higher quantiles. This pattern highlights the importance of targeting exterminator visits to areas experiencing higher rodent activity, where the intervention appears to have the most significant impact. This insight can help NYC optimize its pest control resources, ensuring that heavily impacted neighborhoods receive priority attention. By focusing on high-complaint regions, the city can better address rodent infestations and improve the effectiveness of its rodent management strategies.

## Recommendations

* Target High-Complaint Areas:
Given the strong relationship between exterminator visits and complaints at high quantiles, resources should focus on high-complaint areas to address rodent issues proactively.
  
* Enhance Inspection Effectiveness:
Strengthen compliance inspection processes in areas with high complaints to validate the apparent (though not significant) reduction in complaints associated with inspections.





